---
title: "In-class Ex 1"
date: "18 Nov 23"
date-modified: "last-modified" 

format: html
execute: 
  echo: true
  eval: true
  warning: false
  
editor: visual 
---

## Getting Started

Using pacman to load the required packages.

-   tmap - for thematic mapping

-   sf - for geospatial data handling

-   tidyverse - for non-spatial data handling

```{r}

pacman::p_load(sf, tmap, tidyverse)
```

## Importing kml file

Using the data from LTA on the Location of the Bus Stops in Singapore

Available at: <https://datamall.lta.gov.sg/>

Current data is based on Jul 23 (updated quarterly).

```{r}

busstops = st_read(dsn = "data/geospatial", 
               layer = "BusStop") %>%
  st_transform (crs = 3414)
```

```{r}
plot(busstops)
```

## Bus Origin and Destination Data

We will also import the *Passengers, Origin and Destination* data by using the *readr package* and *read_csv*.

```{r}

odbus_aug <- read_csv("data/aspatial/origin_destination_bus_202308.csv")

odbus_sep <- read_csv("data/aspatial/origin_destination_bus_202309.csv")

odbus_oct <- read_csv("data/aspatial/origin_destination_bus_202310.csv")

head(odbus_aug, n=5)
```

We need to convert the Origin_PT_Code data into factor to allow R to manipulate the data more easily.

```{r}

odbus_aug$ORIGIN_PT_CODE <- as.factor(odbus_aug$ORIGIN_PT_CODE)

odbus_aug$DESTINATION_PT_CODE <- as.factor(odbus_aug$DESTINATION_PT_CODE)
```

Filter the time between 7 am to 9 am (just before 10am) on Weekdays, Group by the Origin Bus Stops.

```{r}

origtrip_7_9 <- odbus_aug %>%
  filter(DAY_TYPE == "WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 7 &
           TIME_PER_HOUR <= 9) %>%
  group_by(ORIGIN_PT_CODE) %>%
  summarise (TRIPS = sum(TOTAL_TRIPS))
```

## Map Planning Sub-zone 2019

We will also import that Map Planning Sub-zone 2019 data. Need to transform the coordinates to 3414, which will be in metres instead of decimal/degrees.

```{r}

mpsz19 = st_read(dsn = "data/geospatial", 
               layer = "MPSZ-2019") %>%
  st_transform (crs = 3414)

head(mpsz19, n=5)

plot(mpsz19)
```

## Joining the data of the Bus Stop Code, Location and which Planning Sub-zone 
