---
title: "Take Home Ex 2"

date: "6 Dec 23"
date-modified: "last-modified" 

format: html
execute: 
  echo: true
  eval: true
  warning: false
  
editor: visual 
---

# Overview of the Exercise

::: panel-tabset
## Setting the Scene

What are the driving forces behind urban dwellers to weak up early in morning to commute from their home locations to their work places? What are the impact of removing a public bus service on the commuters reside along the corridor of the bus route? These and many other questions related to urban mobility are challenges faced by transport operators and urban managers.

To provide answer to this question, traditionally, commuters survey will be used. However, commuters survey is a very costly, time-consuming and laborous, not to mention that the survey data tend to take a long time to clean and analyse. As a result, it is not unusual, by the time the survey report was ready, most of the information already out-of-date!

As city-wide urban infrastructures such as public buses, mass rapid transits, public utilities and roads become digital, the data sets obtained can be used as a framework for tracking movement patterns through space and time. This is particularly true with the recent trend of massive deployment of pervasive computing technologies such as GPS on the vehicles and SMART cards used by public transport commuters.

Unfortunately, this explosive growth of geospatially-referenced data has far outpaced the planner's ability to utilize and transform the data into insightful information thus creating an adverse impact on the return on the investment made to collect and manage this data.

## Motivation and Objective

This take-home exercise is motivated by two main reasons. Firstly, despite increasing amounts of open data available for public consumption, there has not been significant practice research carried out to show how these disparate data sources can be integrated, analysed, and modelled to support policy making decisions.

Secondly, there is a general lack of practical research to show how geospatial data science and analysis (GDSA) can be used to support decision-making.

Hence, your task for this take-home exercise is to conduct a case study to demonstrate the potential value of GDSA to integrate publicly available data from multiple sources for building a spatial interaction models to determine factors affecting urban mobility patterns of public bus transit.

## The Data

### Open Gov Data

For the purpose of this assignment, data from several open government sources will be used:

-   *Passenger Volume by Origin Destination Bus Stops*, *Bus Stop Location*, *Train Station* and *Train Station Exit Point*, just to name a few of them, from [LTA DataMall](https://datamall.lta.gov.sg/content/datamall/en.html).

-   *Master Plan 2019 Subzone Boundary*, *HDB Property Information*, *School Directory and Information* and other relevant data from [Data.gov.sg](https://beta.data.gov.sg/).

-   Added the data of childcare services obtained from onemap.gov.sg

### **Specially collected data**

-   Businesses, retail and services, leisure and recreation, etc geospatial data sets assemble by course instructor. (Refer to eLearn)

## Task

The specific tasks of this take-home exercise are as follows:

Derive an analytical hexagon data of 375m (this distance is the perpendicular distance between the centre of the hexagon and its edges) to represent the [traffic analysis zone (TAZ)](https://tmg.utoronto.ca/files/Reports/Traffic-Zone-Guidance_March-2021_Final.pdf).

-   With reference to the time intervals provided in the table below, construct an O-D matrix of commuter flows for a time interval of your choice by integrating *Passenger Volume by Origin Destination Bus Stops* and *Bus Stop Location* from [LTA DataMall](https://datamall.lta.gov.sg/content/datamall/en.html). The O-D matrix must be aggregated at the analytics hexagon level

    | Peak hour period             | Bus tap on time  | What it Means (Likely)                                            |
    |------------------|------------------|-----------------------------------|
    | ***Weekday morning peak***   | ***6am to 9am*** | ***Going to work and school (our data will be focused on this)*** |
    | Weekday afternoon peak       | 5pm to 8pm       | Returning home from work or school                                |
    | Weekend/holiday morning peak | 11am to 2pm      | Heading out for retail shopping, attractions etc                  |
    | Weekend/holiday evening peak | 4pm to 7pm       | Returning home for Mon Blues                                      |

-   Display the O-D flows of the passenger trips by using appropriate geovisualisation methods (not more than 5 maps).

-   Describe the spatial patterns revealed by the geovisualisation (not more than 100 words per visual).

-   Assemble at least three propulsive and three attractiveness variables by using aspatial and geospatial from publicly available sources.

-   Compute a distance matrix by using the analytical hexagon data derived earlier.

Spatial Interaction Modelling

-   Calibrate spatial interactive models to determine factors affecting urban commuting flows at the selected time interval.

-   Present the modelling results by using appropriate geovisualisation and graphical visualisation methods. (Not more than 5 visuals)

-   With reference to the Spatial Interaction Model output tables, maps and data visualisation prepared, describe the modelling results. (not more than 100 words per visual).
:::

# Getting Started

So for the purpose of this exercise - I will focus on the **Weekdays Morning Travel Period**.

I will also utilise the writing and loading of rds to reduce the rendering and loading time as well as make the site more reproducible.

## Loading the Packages

```{r}
pacman::p_load(tmap, sf, sp, DT,
               performance, reshape2,
               ggpubr, tidyverse, stplanr, jsonlite)

```

## Preparing the Flow Data

::: panel-tabset
## Importing the OD Data

```{r}
#|eval: false

odbus <- read_csv("data/aspatial/origin_destination_bus_202310.csv")
```

## Extracting the Data according to the Weekday AM period

```{r}
#|eval: false
odbus6_9 <- odbus %>%
  filter(DAY_TYPE == "WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 6 &
           TIME_PER_HOUR <= 9) %>%
  group_by(ORIGIN_PT_CODE,
           DESTINATION_PT_CODE) %>%
  summarise(TRIPS = sum(TOTAL_TRIPS))


```

## Working with Geospatial Data

```{r}
#|eval: false
busstops = st_read(dsn = "data/geospatial", 
               layer = "BusStop") %>%
  st_transform (crs = 3414)
```

## Creating Hexagonal Grids

Create analytical hexagon data of 325m (this distance is the perpendicular distance between the centre of the hexagon and its edges) - so should be **c(750, 750)**.

```{r}
#|eval: false
hexagon_grid = st_make_grid(busstops, c(750, 750), what = "polygons", square = FALSE)

hexagon_grid_sf = st_sf(hexagon_grid) %>%
  mutate(grid_id = 1:length(lengths(hexagon_grid)))
```

## Combine with Hexagon Grid

```{r}
#|eval: false
hexagon_grid_sf$n_busstops = lengths(st_intersects(hexagon_grid_sf, busstops))

hexagon_busstop = filter(hexagon_grid_sf, n_busstops > 0)

busstop_grid <- st_intersection(busstops, hexagon_grid_sf)%>%
  select(BUS_STOP_N, grid_id) %>%
  st_drop_geometry()



```

## Appending Data and Checking

```{r}
#|eval: false
od_data_wdam <- left_join(odbus6_9 , busstop_grid,
            by = c("ORIGIN_PT_CODE" = "BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_Grid = grid_id,
         DESTIN_BS = DESTINATION_PT_CODE)
```

```{r}
#|eval: false
duplicate <- od_data_wdam %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```

```{r}
#|eval: false
od_data_wdam <- unique(od_data_wdam)

```

```{r}
#|eval: false
od_data_wdam <- left_join(od_data_wdam , busstop_grid,
            by = c("DESTIN_BS" = "BUS_STOP_N")) 

duplicate <- od_data_wdam %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()

od_data_wdam <- unique(od_data_wdam)

od_data_wdam <- od_data_wdam %>%
  rename(DESTIN_Grid = grid_id) %>%
  drop_na() %>%
  group_by(ORIGIN_Grid, DESTIN_Grid) %>%
  summarise(MORNING_PEAK = sum(TRIPS))

#Removing Intra-Grid Flow
od_wdam_cleaned <- od_data_wdam[od_data_wdam$ORIGIN_Grid!=od_data_wdam$DESTIN_Grid,]
```
:::

## Write/Read Data into rds for Subsequent Usage

```{r}
#| code-fold: true
write_rds(hexagon_busstop, "data/rds/hexagon_busstop.rds")

write_rds(od_wdam_cleaned, "data/rds/od_wdam_cleaned.rds")

write_rds(hexagon_grid_sf, "data/rds/hexagon_grid_sf.rds")
```

```{r}
#| code-fold: true

hexagon_bustops <- read_rds("data/rds/hexagon_busstop.rds")

od_wdam_cleaned <- read_rds("data/rds/od_wdam_cleaned.rds")

hexagon_grid_sf <- read_rds("data/rds/hexagon_grid_sf.rds")

schools_sf <- read_rds("data/rds/schools_sf.rds")

mpsz <- read_rds("data/rds/mpsz.rds")

flow_data_tidy <- read_rds("data/rds/flow_data_tidy.rds")

```

# Visualising the Spatial Interaction

```{r}
#| code-fold: true
#| eval: false

flowLine <- od2line(flow = od_wdam_cleaned, 
                    zones = hexagon_grid_sf,
                    zone_code = "grid_id")
```

```{r}
#| code-fold: true

tmap_options(check.and.fix = TRUE)
tmap_mode("plot")

tm_shape(mpsz) +
  tm_polygons(col = "black", alpha = 0.5) +
  tm_layout(main.title = "Spatial Flow for Passengers by Bus on Weekdays Morning",
            main.title.position = "center",
            main.title.size = 1.0,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2, labels.show = FALSE)+
tm_shape(hexagon_bustops) +
  tm_polygons(alpha = 0.3, col = "lightblue") +
flow_data_tidy %>%  
  filter(TRIPS >= 5000) %>%
tm_shape() +
  tm_lines(lwd = "TRIPS",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           col = "yellow",
           alpha = 0.5)
```

```{r}
#| code-fold: true

tmap_options(check.and.fix = TRUE)
tmap_mode("plot")

tm_shape(mpsz) +
  tm_polygons(col = "black", alpha = 0.5) +
  tm_layout(main.title = "Spatial Flow for Passengers by Bus on Weekdays Morning, Dist > 10km",
            main.title.position = "center",
            main.title.size = 1.0,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2, labels.show = FALSE)+
tm_shape(hexagon_bustops) +
  tm_polygons(alpha = 0.3, col = "lightblue") +
flow_data_tidy %>%  
  filter(TRIPS>=5000, DIST >= 10000) %>%
tm_shape() +
  tm_lines(lwd = "TRIPS",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           col = "yellow",
           alpha = 0.5)
```

# Analysis of Spatial Flow

From the spatial flow, it seems that there are a few identified areas where people are taking buses to during the morning peak period:

-   Travelling from Malaysia to Singapore through the Causeway - which accounts for a very high number. And it seems they will make their way towards Woodlands and branch out from that region.

-   Other regions include areas like Yishun, Tampines, Jurong, Boon Lay.

-   There are some long distances travel between Woodlands towards Changi, i.e. from the North to the Eastern Part of Singapore. This is probably due to a lack of direct MRT Lines, and that the Buses travelling on the Expressway would provide a speedier route.

# Spatial Patterns

## Attractiveness Variables

We will compile three attractiveness variables, namely:

-   **Schools** (their place of study)

-   **Business** (their place of work)

-   **F&B** (to buy breakfast before heading to their final destination, assuming that they will buy food nearer to their location)

-   **Financial Services** (locations about the banks and money changers in Singapore - assuming that people will travel there to work in the morning)

::: panel-tabset
## School Data (from Inclass Ex 4)

```{r}
#| eval: false

hexagon_bustops$SCHOOL_COUNT <- lengths(
  st_intersects(
    hexagon_bustops, schools_sf))
```

```{r}
#| eval: false

sum(hexagon_bustops$SCHOOL_COUNT)
```

## Business Data

We assume that the businesses equate to the companies etc that people will travel to for their work.

```{r}
#| eval: false
business_sf <- st_read(dsn = "data/geospatial", 
               layer = "Business")
```

```{r}
#| eval: false
business_sf<- unique(business_sf)
```

```{r}
#| eval: false

hexagon_bustops$BIZ_COUNT <- lengths(
  st_intersects(
    hexagon_bustops, business_sf))
```

## F&B Data

```{r}
#| eval: false

food_sf <- st_read(dsn = "data/geospatial", 
               layer = "F&B") %>%
  st_transform (crs=3414)

```

```{r}
#| eval: false
food_sf<- unique(food_sf)
```

```{r}
#| eval: false

hexagon_bustops$FOOD_COUNT <- lengths(
  st_intersects(hexagon_bustops, food_sf))
```

## Financial Services Data

```{r}
#| eval: false

finserv_sf <- st_read(dsn='data/geospatial',
                  layer='FinServ') %>% 
  st_transform(crs=3414)
```

```{r}
#| eval: false
finserv <- unique(finserv_sf)
```

```{r}
#| eval: false

hexagon_bustops$FINSERV_COUNT <- lengths(
  st_intersects(hexagon_bustops, finserv_sf))
```
:::

## Propulsive Variables

We will look at three propulsive variables, namely:

-   Number of Population

-   Number of Train Stations Exits

-   Number of Bus Stops (already available since Take Home Ex 1)

-   Location of Pre Schools (assuming parents will bring their child to the nearest pre-school which should be within walking distance, before going to the nearest bus stop to travel for work)

-   Number of Buses available within each grid - tried to obtain the bus route data from LTA Data Mall, but there data seems limited and outdated (no actual date on when it was last updated); so decided not to pursue this.

### Preparing the Data on Populace

We utilise the hdb data provided by Prof.

```{r}
#| eval: false
hdb <- read_csv ("data/aspatial/hdb.csv")
```

Based on the data, we will be able to intersect the locations, types of dwellings (whether 1-room, 2-room, rental and etc) to have an estimate on the number of population in each grid.

So based on past studies in 2018:

| Types of Dwelling | Average Household Size | Remarks |
|-------------------|------------------------|---------|
| 1-Room / Rental   | 2.3                    |         |
| 2-Room            | 2.19                   |         |
| 3-Room            | 2.63                   |         |
| 4-Room            | 3.42                   |         |
| 5-Room / Exec     | 3.77                   |         |

```{r}
#| eval: false
hdb_tidy <- hdb %>%
  select(13:25, "lat", "lng") %>%
  mutate(one_Rm = (`1room_sold` + `1room_rental` + studio_apartment_sold + other_room_rental)*2.3) %>%
  mutate(two_Rm = (`2room_sold` + `2room_rental`)*2.19) %>%
  mutate(three_Rm = (`3room_sold` + `3room_rental`)*2.63) %>%
  mutate(four_Rm = (`4room_sold`) * 3.42) %>%
  mutate(five_rm = (`5room_sold` + `exec_sold` + multigen_sold)*3.77)
```

```{r}
#| eval: false
hdb_tidy <- hdb_tidy %>%
  select (14:20) %>%
  mutate (pop = one_Rm + two_Rm + three_Rm + four_Rm + five_rm)
```

Covert to sf data.

```{r}
#| eval: false
hdb_sf <- st_as_sf(hdb_tidy,
                      coords = c("lng", "lat"),
                      crs = 4326) %>%
  st_transform (crs=3414)

write_rds (hdb_sf, "data/rds/hdb_sf.rds")
```

```{r}
#| eval: false
hexagon_bustops$POP <- sapply(
  st_intersects(hexagon_bustops, hdb_sf),
  function(indices) sum(hdb_sf$pop[indices])
)
```

### Train Station Exit

I think that the train station exits can be considered as both propulsive and attractive, e.g. people will take buses to the nearest train station exit, then take on the mrt to their next destination.

But they can also exit from the train station, then proceed to take bus to their place of work or study.

I downloaded the static shp file of the train station exits from LTADataMall.

```{r}
#| eval: false
train_exits_sf <- st_read(dsn = "data/geospatial", 
               layer = "Train_Station_Exit_Layer") %>%
  st_transform (crs=3414)
```

```{r}
#| eval: false
hexagon_bustops$TRAIN_EXITS_COUNT <- lengths(
  st_intersects(
    hexagon_bustops, train_exits_sf))
```

### Location of Pre-Schools

Using the Onemap API, I gain access to the "Child Care Services" theme.

This is updated by ECDA in Dec 23 and i saved it in JSON format.

I will use the `jsonlite` package to convert the data.

```{r}
#| eval: false
childcare <- fromJSON("data/aspatial/childcare.json", flatten = TRUE)
```

```{r}
#| eval: false
childcare_tidy <- childcare %>%
  select(9:14) %>%
  slice(-1) %>%
  separate(col = SrchResults.LatLng, into = c("lat", "long"), sep = ",", convert = TRUE)
```

```{r}
#| eval: false

childcare_sf <- st_as_sf(childcare_tidy,
                      coords = c("long", "lat"),
                      crs = 4326) %>%
  st_transform (crs=3414)

```

```{r}
#| eval: false

hexagon_bustops$CHILDCARE_COUNT <- lengths(
  st_intersects(
    hexagon_bustops, childcare_sf))
```

All the data were combined and written into an rds file.

```{r}
#| eval: false
write_rds (hexagon_bustops, "data/rds/combined_data.rds")
```

```{r}
combined_data <- read_rds("data/rds/combined_data.rds")
```

# Visualising the Flow with the Variables

::: panel-tabset
## With School Count

```{r}
#| code-fold: true

tmap_options(check.and.fix = TRUE)
tmap_mode("plot")

tm_shape(mpsz) +
  tm_polygons(col = "grey", alpha = 0.5) +
  tm_layout(main.title = "Spatial Flow for Passengers by Bus on Weekdays Morning",
            main.title.position = "center",
            main.title.size = 1.0,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2, labels.show = FALSE)+
tm_shape(combined_data[combined_data$SCHOOL_COUNT > 1, ]) +
  tm_borders() +
    tm_fill("SCHOOL_COUNT", alpha = 0.3) +
flow_data_tidy %>%  
  filter(TRIPS>=5000) %>%
tm_shape() +
  tm_lines(lwd = "TRIPS",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           col = "yellow",
           alpha = 0.5)
```

## With Business Count

```{r}
#| code-fold: true

tmap_options(check.and.fix = TRUE)
tmap_mode("plot")

tm_shape(mpsz) +
  tm_polygons(col = "grey", alpha = 0.5) +
  tm_layout(main.title = "Spatial Flow for Passengers by Bus on Weekdays Morning",
            main.title.position = "center",
            main.title.size = 1.0,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2, labels.show = FALSE)+
tm_shape(combined_data[combined_data$SCHOOL_COUNT > 1, ]) +
  tm_borders() +
    tm_fill("BIZ_COUNT", alpha = 0.3) +
flow_data_tidy %>%  
  filter(TRIPS>=5000) %>%
tm_shape() +
  tm_lines(lwd = "TRIPS",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           col = "yellow",
           alpha = 0.5)
```

## With Population

```{r}
#| code-fold: true

tmap_options(check.and.fix = TRUE)
tmap_mode("plot")

tm_shape(mpsz) +
  tm_polygons(col = "grey", alpha = 0.5) +
  tm_layout(main.title = "Spatial Flow for Passengers by Bus on Weekdays Morning",
            main.title.position = "center",
            main.title.size = 1.0,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2, labels.show = FALSE)+
tm_shape(combined_data[combined_data$SCHOOL_COUNT > 1, ]) +
  tm_borders() +
    tm_fill("POP", alpha = 0.3) +
flow_data_tidy %>%  
  filter(TRIPS>=5000) %>%
tm_shape() +
  tm_lines(lwd = "TRIPS",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           col = "yellow",
           alpha = 0.5)
```
:::

# Spatial Interaction Modelling

## Computing the Distance Matrix

```{r}
#| eval: false  

hexa_sp <- as(hexagon_busstop,"Spatial")  
hexa_sp
```

```{r}
#| eval: false 

dist <- spDists(hexa_sp,                  
                longlat = FALSE) #latlong is FALSE cuz its is in Svy21 alr  
head(dist, n=c(10, 10))
```

```{r}
#| eval: false 
hexa_id <- hexagon_busstop$grid_id
```

```{r}
#| eval: false 
colnames(dist) <- paste0(hexa_id) rownames(dist) <- paste0(hexa_id)
```

```{r}
#| eval: false 
distPair <- melt(dist) %>%   rename(dist = value) head(distPair, 10)
```

```{r}
#| eval: false 

distPair$dist <- ifelse(distPair$dist == 0,                         
                        50, distPair$dist)
```

```{r}
#| eval: false 
distPair <- distPair %>%   rename(orig = Var1,          
                                  dest = Var2)
```

```{r}
#| eval: false 
write_rds(distPair, "data/rds/distPair.rds") 
```

```{r}
distPair <- read_rds("data/rds/distPair.rds")
```

## Preparing the Flow Data

### Grouping the Flow Data

```{r}
flow_data <- od_wdam_cleaned %>%
  group_by(ORIGIN_Grid, DESTIN_Grid) %>% 
  summarize(TRIPS = sum(MORNING_PEAK)) 
```

```{r}
flow_data <- od2line(flow = flow_data, 
                    zones = hexagon_grid_sf,
                    zone_code = "grid_id")
```

### Separating the Intra-Flow Data

```{r}
flow_data$FlowNoIntra <- ifelse(
  flow_data$ORIGIN_Grid == flow_data$DESTIN_Grid, 
  0, flow_data$TRIPS)
flow_data$offset <- ifelse(
  flow_data$ORIGIN_Grid == flow_data$DESTIN_Grid, 
  0.000001, 1)
```

### Combining Passenger Volume Data with Dist Data

```{r}
flow_data1 <- flow_data %>%
  left_join (distPair,
             by = c("ORIGIN_Grid" = "orig",
                    "DESTIN_Grid" = "dest"))
```

### Combining the Data with Propulsive and Attractive Variables Data

```{r}
#| eval: false
combined_tidy <- combined_data %>%
  st_drop_geometry()
```

```{r}
#| eval: false
combined_tidy$DESTIN_TRAIN_EXITS <- combined_tidy$TRAIN_EXITS_COUNT
```

```{r}
#| eval: false
flow_data_tidy <- flow_data1 %>%
  left_join(
    combined_tidy %>% 
      select(grid_id, SCHOOL_COUNT, BIZ_COUNT, FINSERV_COUNT, FOOD_COUNT, DESTIN_TRAIN_EXITS),
      by = c("DESTIN_Grid" = "grid_id")
  ) %>%
  rename(DIST = dist)
```

```{r}
#| eval: false

flow_data_tidy <- flow_data_tidy %>%
  left_join(
    combined_tidy %>% 
      select(grid_id, POP, TRAIN_EXITS_COUNT, CHILDCARE_COUNT, n_busstops),
      by = c("ORIGIN_Grid" = "grid_id")
  ) 
```

Ensure variables with value = 0 are adjusted to 0.99.

```{r}
#| eval: false
#| code-fold: true

flow_data_tidy$SCHOOL_COUNT <- ifelse(
  flow_data_tidy$SCHOOL_COUNT == 0,
  0.99, flow_data_tidy$SCHOOL_COUNT)

flow_data_tidy$BIZ_COUNT <- ifelse(
  flow_data_tidy$BIZ_COUNT == 0,
  0.99, flow_data_tidy$BIZ_COUNT)

flow_data_tidy$FINSERV_COUNT <- ifelse(
  flow_data_tidy$FINSERV_COUNT == 0,
  0.99, flow_data_tidy$FINSERV_COUNT)

flow_data_tidy$FOOD_COUNT <- ifelse(
  flow_data_tidy$FOOD_COUNT == 0,
  0.99, flow_data_tidy$FOOD_COUNT)

flow_data_tidy$TRAIN_EXITS_COUNT <- ifelse(
  flow_data_tidy$TRAIN_EXITS_COUNT == 0,
  0.99, flow_data_tidy$TRAIN_EXITS_COUNT)

flow_data_tidy$DESTIN_TRAIN_EXITS <- ifelse(
  flow_data_tidy$DESTIN_TRAIN_EXITS == 0,
  0.99, flow_data_tidy$DESTIN_TRAIN_EXITS)

flow_data_tidy$POP <- ifelse(
  flow_data_tidy$POP == 0,
  0.99, flow_data_tidy$POP)

flow_data_tidy$CHILDCARE_COUNT <- ifelse(
  flow_data_tidy$CHILDCARE_COUNT == 0,
  0.99, flow_data_tidy$CHILDCARE_COUNT)

```

```{r}
#| eval: false
write_rds (flow_data_tidy, "data/rds/flow_data_tidy.rds")
```

# Calibrating Spatial Interaction Models

```{r}
flow_data_tidy <- read_rds ("data/rds/flow_data_tidy.rds")
```

We will calibrate separate Spatial Interaction Models for inter- and intra-zonal flows. For the Ex, we will focus our attention on inter-zonal flow. Hence, we need to exclude the intra-zonal flow from *flow_data*.

First, two new columns called *FlowNoIntra* and *offset* will be created by using the code chunk below.

```{r}
flow_data_tidy$FlowNoIntra <- ifelse(
  flow_data_tidy$ORIGIN_Grid == flow_data_tidy$DESTIN_Grid, 
  0, flow_data_tidy$TRIPS)

flow_data_tidy$offset <- ifelse(
  flow_data_tidy$ORIGIN_Grid == flow_data_tidy$DESTIN_Grid, 
  0.000001, 1)
```

According to the syntax used to derive values in *FlowNoIntra* field, all intra-zonal flow will be given a value of 0 or else the original flow values will be inserted.

Next, inter-zonal flow will be selected from flow_data and save into a new output data.frame called *inter_zonal_flow* by using the code chunk below.

```{r}
inter_zonal_flow <- flow_data_tidy %>%
  filter(FlowNoIntra > 0) %>%
  mutate(ORIGIN_Grid = as.factor(ORIGIN_Grid)) %>%
  mutate(DESTIN_Grid = as.factor(DESTIN_Grid)) 
```

## Basic Plot of the Inter Zonal Flow Data

```{r}
ggplot(data = inter_zonal_flow,
       aes(x = TRIPS)) +
  geom_histogram(color='black',size= 0.3, fill = 'lightblue') +
  labs(y= 'Count of TRIPS', x='TRIPS') +
  theme(axis.title.y=element_text(angle = 0),
        plot.title = element_text(face="bold")) +
  ggtitle('Distribution of TRIPS')
```

It seems to suggest that we should use a **Poisson Regression Model**.

## Unconstrained Spatial Interaction Model

```{r}
#| eval: false
uncSIM <- glm(formula = TRIPS ~ 
                log(SCHOOL_COUNT) +
                log(BIZ_COUNT) +
                log(FINSERV_COUNT) +
                log(FOOD_COUNT) +
                log(POP) +
                log(TRAIN_EXITS_COUNT) +
                log(DESTIN_TRAIN_EXITS) +
                log(CHILDCARE_COUNT) +
                log(n_busstops) +
                log(DIST),
              family = poisson(link = "log"),
              data = inter_zonal_flow,
              na.action = na.exclude)

write_rds(uncSIM, "data/rds/uncSIM.rds")

```

```{r}
uncSIM <- read_rds("data/rds/uncSIM.rds")
```

```{r}
#| eval: false
summary(uncSIM)
```

```{r}
uncSIM$coefficients
```

Based on the results, all the coefficients are statistically significant. The most influential explanatory variables areÂ *distance*,Â *destination train exit count*Â andÂ *origin number of bus stops* with coefficients of -1.45, 0.475 and 0.295 respectively.

## Origin - Constrained Model

::: callout-note
-   For origin-constrained model, only explanatory variables representing the attractiveness at the destinations will be used.

-   All the explanatory variables including distance will be log transformed.

-   *ORIGIN_Grid* is used to model ðœ‡~ð‘–~ . It must be in categorical data type.

-   It is important to note that -1 is added in the equation after the distance variable. The -1 serves the purpose of removing the intercept that by default, glm will insert into the model.
:::

```{r}
#| eval: false
orcSIM <- glm(formula = TRIPS ~ 
                ORIGIN_Grid +
                log(SCHOOL_COUNT) +
                log(BIZ_COUNT) +
                log(FINSERV_COUNT) +
                log(FOOD_COUNT) +
                log(DESTIN_TRAIN_EXITS) +
                log(DIST) - 1,
              family = poisson(link = "log"),
              data = inter_zonal_flow,
              na.action = na.exclude)

write_rds(orcSIM, "data/rds/orcSIM.rds")
```

```{r}

orcSIM <- read_rds("data/rds/orcSIM.rds")
```

```{r}
#| eval: false
tail(summary(orcSIM)$coefficients, n = 10)
```

Based on the results, all of the coefficients are statistically significant. The most influential explanatory variables areÂ *distance*,Â *destination train exit count*Â andÂ *destination school count* with coefficients of -1.5, 0.8 and 0.22 respectively.

## Destination Constrained Model

Look at the propulsive variables.

```{r}
#| eval: false
decSIM <- glm(formula = TRIPS ~ 
                DESTIN_Grid +
                log(POP) +
                log(CHILDCARE_COUNT) +
                log(TRAIN_EXITS_COUNT) +
                log(n_busstops) +
                log(DIST) - 1,
              family = poisson(link = "log"),
              data = inter_zonal_flow,
              na.action = na.exclude)

write_rds(decSIM, "data/rds/decSIM.rds")
```

```{r}

decSIM <- read_rds("data/rds/decSIM.rds")
```

```{r}
tail(summary(decSIM)$coefficients, n = 10)
```

Based on the results, all of the coefficients are statistically significant. The most influential explanatory variables areÂ *distance*,Â *origin number of bus stops*Â andÂ *origin train exit count* with coefficients of -1.48, 0.29 and 0.285 respectively.

## Doubly Constrained Model

```{r}
#| eval: false
dbcSIM <- glm(formula = TRIPS ~ 
                ORIGIN_Grid + 
                DESTIN_Grid + 
                log(DIST),
              family = poisson(link = "log"),
              data = inter_zonal_flow,
              na.action = na.exclude)

```

```{r}
#| eval: false
write_rds (dbcSIM, "data/rds/dbcSIM.rds")
```

```{r}
dbcSIM <- read_rds("data/rds/dbcSIM.rds")
```

```{r}
tail(summary(dbcSIM)$coefficients, n = 10)
```

The variable *distance* is statistically significant.

# Model Comparisons

We will use [`compare_performance()`](https://easystats.github.io/performance/reference/compare_performance.html) of [**performance**](https://easystats.github.io/performance/) package.

First of all, let us create a list called *model_list* by using the code chunk below.

```{r}
model_list <- list(unconstrained=uncSIM,
                   originConstrained=orcSIM,
                   destinationConstrained=decSIM,
                   doublyConstrained=dbcSIM)
```

Next, we will compute the RMSE of all the models inÂ *model_list*Â file by using the code chunk below.

```{r}
compare_performance(model_list,
                    metrics = "RMSE")
```

## Goodness of Fit - Calculating RSquared Statistics

```{r}
CalcRSquared <- function(observed, estimated){
  r <- cor(observed, estimated)
  R2 <- r^2
  R2
}
```

```{r}

r_squared_values <- lapply(model_list, function(model) {
  observed_values <- model$data$TRIPS
  fitted_values <- model$fitted.values
  CalcRSquared(observed_values, fitted_values)
})

# Print the R-squared values for each model
print(r_squared_values)

```

It seems that the Doubly Constrained Model is the best fitted across the four models, as compared to the other three. the RMSE of the Doubly Constrained Model is also the lowest.

## Visualising the Fitted Values

```{r}
#| code-fold: true
df <- as.data.frame(uncSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
#| code-fold: true
SIM_data <- inter_zonal_flow %>%
  cbind(df) %>%
  rename(uncTRIPS = "uncSIM.fitted.values")
```

```{r}
#| code-fold: true
df <- as.data.frame(orcSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
#| code-fold: true
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(orcTRIPS = "orcSIM.fitted.values")
```

```{r}
#| code-fold: true
df <- as.data.frame(decSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
#| code-fold: true
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(decTRIPS = "decSIM.fitted.values")
```

```{r}
#| code-fold: true
df <- as.data.frame(dbcSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
#| code-fold: true
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(dbcTRIPS = "dbcSIM.fitted.values")
```

```{r}
write_rds(SIM_data, "data/rds/SIM_data.rds")
```

```{r}
#| code-fold: true
unc_p <- ggplot(data = SIM_data,
                aes(x = uncTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

orc_p <- ggplot(data = SIM_data,
                aes(x = orcTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

dec_p <- ggplot(data = SIM_data,
                aes(x = decTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

dbc_p <- ggplot(data = SIM_data,
                aes(x = dbcTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

ggarrange(unc_p, orc_p, dec_p, dbc_p,
          ncol = 2,
          nrow = 2)
```

From the graphs - I noticed that as the number of trips increase, the fit is not as good. Additionally, there are quite a fair bit of variability in the data.

# Limitations about Spatial Interaction Model

I realised that there are some limitations to the Spatial Interaction Models, namely:

-   The models are **sensitive to distance** - distance seems to be still the most influential factor in predicting the flow.

-   The models ignore the **socio-economic factors** of the population.

-   The models **fail to explain the complex spatial pattern**s, e.g. accounting for the dependence between the hexagons, i.e. they are not independent.

# **Spatial Econometric Interaction Modelling (SEIM)**

Since Prof talked about the SEIM in the class today, I want to see how we can use it for this take home ex 2.

## Package Required

In this exercise, the development version (0.1.0.9010) ofÂ **spflow**Â will be used instead of the released version (0.1.0). The code chunk below will be used to install the development version ofÂ **spflow**Â package.

```{r}
devtools::install_github("LukeCe/spflow")
```

```{r}
pacman::p_load(tmap, sf, spdep, sp, Matrix,
               spflow, reshape2, knitr,
               tidyverse)
```

## Data Preparation

Before we can calibrate Spatial Econometric Interaction Models by using **spflow** package, three data sets are required. They are:

-   a spatial weights,

-   a tibble data.frame consists of the origins, destination, flows and distances between the origins and destination, and

-   a tibble data.frame consists of the explanatory variables.

Our geographical data has already been built in the earlier segments and using the hexagonal grid.

### Preparing the Weights Data

```{r}
#| eval: false
grid_busstops <- read_rds("data/rds/hexagon_busstop.rds")
```

```{r}
centroids <- suppressWarnings({
    st_point_on_surface(st_geometry(hexagon_busstop))})

grid_nb <- list(
  "by_contiguity" = poly2nb(hexagon_busstop),
  "by_distance" = dnearneigh(centroids, 
                             d1 = 0, d2 = 5000),
  "by_knn" = knn2nb(knearneigh(centroids, 3))
)
```

Realised that there were data with zero neighbours. 3 data were removed.

```{r}
#| eval: false
nb_contiguity <- grid_nb$by_contiguity

nonzero_indices <- which(card(nb_contiguity) > 0)

hexagon_busstop <- hexagon_busstop %>% slice(nonzero_indices)

```

```{r}
#| eval: false
write_rds(grid_nb, "data/rds/grid_nb.rds")
```

### Preparing the Flow Data

```{r}
#| eval: false
busstop_grid <- st_intersection(busstops, hexagon_grid_sf) %>%
  select(BUS_STOP_N, grid_id) %>%
  st_drop_geometry()
```

```{r}
#| eval: false
od_data <- left_join(odbus6_9 , busstop_grid,
            by = c("ORIGIN_PT_CODE" = "BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_GRID = grid_id,
         DESTIN_BS = DESTINATION_PT_CODE)
```

```{r}
#| eval: false
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```

```{r}
#| eval: false
od_data <- unique(od_data)
```

```{r}
#| eval: false
od_data <- left_join(od_data , busstop_grid,
            by = c("DESTIN_BS" = "BUS_STOP_N")) 
```

```{r}
#| eval: false
od_data <- od_data %>%
  rename(DESTIN_GRID = grid_id) %>%
  drop_na() %>%
  group_by(ORIGIN_GRID, DESTIN_GRID) %>%
  summarise(TRIPS = sum(TRIPS))
```

```{r}
#| eval: false
write_rds(od_data, "data/rds/od_data.rds")
```

```{r}
od_data <- read_rds("data/rds/od_data.rds")
```

```{r}
#| eval: false
busstop_grid_sp <- as(hexagon_busstop, "Spatial")
```

```{r}
#| eval: false
DISTANCE <- spDists(busstop_grid_sp, 
                    longlat = FALSE)
```

```{r}
#| eval: false
grid_names <- hexagon_busstop$grid_id
```

```{r}
#| eval: false
colnames(DISTANCE) <- paste0(grid_names)
rownames(DISTANCE) <- paste0(grid_names)
```

```{r}
#| eval: false
distPair <- melt(DISTANCE) %>%
  rename(DISTANCE = value)
head(distPair, 10)
```

```{r}
#| eval: false
distPair <- distPair %>%
  rename(ORIGIN_GRID = Var1,
         DESTIN_GRID = Var2)
```

```{r}
#| eval: false
flow_data <- distPair %>%
  left_join (od_data) %>%
  mutate(TRIPS = coalesce(TRIPS, 0))
```

```{r}
#| eval: false
write_rds(flow_data, "data/rds/grid_flow.rds")
```

### Preparing the Variables Data

Using the data prepared in the earlier segments. But removing those with zero neighbours.

```{r}
#| eval: false
grid_var <- combined_data

grid_var <- grid_var %>% slice(nonzero_indices)
```

```{r}
#| eval: false
write_rds(grid_var, "data/rds/grid_var.rds")
```

## Preparing spflow objects

```{r}
grid_nb <- read_rds("data/rds/grid_nb.rds")
grid_flow <- read_rds("data/rds/grid_flow.rds")
grid_var <- read_rds("data/rds/grid_var.rds")
```

Three **spflow** objects are required, they are:

-   [`spflow_network-class`](https://lukece.github.io/spflow/reference/spflow_network-class.html), an S4 class that contains all information on a spatial network which is composed by a set of nodes that are linked by some neighborhood relation.

-   [`spflow_network_pair-class`](https://lukece.github.io/spflow/reference/spflow_network_pair-class.html), an S4 class which holds information on origin-destination (OD) pairs. Each OD pair is composed of two nodes, each belonging to one network. All origin nodes must belong to the same origin network should be contained in one `spflow_network-class`, and likewise for the destinations.

-   [`spflow_network_multi-class`](https://lukece.github.io/spflow/reference/spflow_network_multi-class.html), an S4 class that gathers information on multiple objects of types `spflow_network-class` and `spflow_network_pair-class`. Its purpose is to ensure that the identification between the nodes that serve as origins or destinations, and the OD-pairs is consistent (similar to relational data bases).

### **Creating `spflow_network-class` objects**

`spflow_network-class` is an S4 class that contains all information on a spatial network which is composed by a set of nodes that are linked by some neighborhood relation. It can be created by using [`spflow_network()`](https://lukece.github.io/spflow/reference/spflow_network.html) of **spflow** package.

For our model, we choose the contiguity based neighborhood structure.

```{r}
grid_net <- spflow_network(
  id_net = "sg",
  node_neighborhood = nb2mat(grid_nb$by_contiguity),
  node_data = grid_var,
  node_key_column = "grid_id")

grid_net
```

### **Creating `spflow_network-class` object**

`spflow_network-class` object is an S4 class which holds information on origin-destination (OD) pairs. Each OD pair is composed of two nodes, each belonging to one network. All origin nodes must belong to the same origin network should be contained in one `spflow_network-class` object and likewise for the destinations.

In **spflow** package, [`spflow_network_pair()`](https://lukece.github.io/spflow/reference/spflow_network_pair.html)`.`

```{r}
grid_net_pairs <- spflow_network_pair(
  id_orig_net = "sg",
  id_dest_net = "sg",
  pair_data = grid_flow,
  orig_key_column = "ORIGIN_GRID",
  dest_key_column = "DESTIN_GRID")

grid_net_pairs
```

### **Creating `spflow_network_multi-class` object**

The sp_multi_network-class combines information on the nodes and the node-pairs and also ensures that both data sources are consistent. For example, if some of the origins in the sp_network_pair-class are not identified with the nodes in the sp_network_nodes-class an error will be raised.

[`spflow_network_multi()`](https://lukece.github.io/spflow/reference/spflow_network_multi.html)

```{r}
grid_multi_net <-  spflow_network_multi(grid_net,
                                  grid_net_pairs)
grid_multi_net
```

```{r}
#| eval: false

plot(mpsz$geometry)

spflow_map(
  grid_multi_net,
  flow_var = "TRIPS",
  add = TRUE, 
  legend_position = "bottomleft",
  filter_lowest = .99999, 
  remove_intra = TRUE,
  cex = 1)
```

## Model Calibration

The core function of the package is `spflow()`. It provides an interface to three different estimators of spatial econometric interaction models (Dargel 2021) that allow the user to estimate origin-destination flows with spatial autocorrelation.

The three different estimators currently supported by spflow are:

-   **Maximum Likelihood Estimation (MLE)** which is the default estimation procedure. The matrix form estimation in the framework of this model was first developed by LeSage and Pace (2008) and then improved by Dargel (2021) . Spatial two-stage least squares (S2SLS)

-   **Spatial Two-stage Least Squares (S2SLS)** estimator is an adaptation of the one proposed by Kelejian and Prucha (1998), to the case of origin-destination flows, with up to three neighborhood matrices Dargel (2021). A similar estimation is done by Tamesue and Tsutsumi (2016). The user can activate the S2SLS estimation via the estimation_control argument using the input spflow_control(estimation_method = \"s2sls\").

-   **Bayesian Markov Chain Monte Carlo (MCMC)** estimator is based on the ideas of LeSage and Pace (2009) and incorporates the improvements proposed in Dargel (2021) . The estimation is based on a tuned Metropolis-Hastings sampler for the auto-regressive parameters, and for the remaining parameters it uses Gibbs sampling. The routine uses 5500 iterations of the sampling procedure and considers the first 2500 as burn-in period. The user can activate the S2SLS estimation via the estimation_control argument using the input spflow_control(estimation_method = \"mcmc\").

Estimation with default settings requires two arguments: an **sp_multi_network-class** and a **flow_formula**. The flow_formula specifies the model we want to estimate. The function offers a formula interface adapted to spatial interaction models, which has the following structure: Y \~ O\_(X1) + D\_(X2) + I\_(X3) + P\_(X4). This structure reflects the different data sources involved in such a model. On the left hand side there is the independent variable Y which corresponds to the vector of flows. On the right hand side we have all the explanatory variables. The functions O\_(...) and D\_(...) indicate which variables are used as characteristics of the origins and destinations respectively. Similarly, I\_(...) indicates variables that should be used for the intra-regional parameters. Finally, P\_(...) declares which variables describe origin-destination pairs, which most frequently will include a measure of distance.

All the declared variables must be available in the provided `spflow_network_multi()` object, which gathers information on the origins and destinations (inside `spflow_network()` objects), as well as the information on the origin-destination pairs (inside a `spflow_network_pair()` object).

Using the short notation Y \~ . is possible and will be interpreted as usual, in the sense that we use all variables that are available for each data source. Also mixed formulas, such as Y \~ . + P\_(log(X4) + 1), are possible. When the dot shortcut is combined with explicit declaration, it will only be used for the non declared data sources.

### Using Model Control

```{r}
spflow_formula <- log(1 + TRIPS) ~ 
    O_(n_busstops +
    TRAIN_EXITS_COUNT + 
    FOOD_COUNT +
    POP +
    CHILDCARE_COUNT) +
    D_(SCHOOL_COUNT +
    BIZ_COUNT +
    FINSERV_COUNT +
    n_busstops) +
    P_(log(DISTANCE + 1))

model_control <- spflow_control(
  estimation_method = "mle",
  model = "model_8")

mle_model8 <- spflow(
  spflow_formula,
  spflow_networks = grid_multi_net,
  estimation_control = model_control)

mle_model8
```

The R2 is 0.718 which is higher compared to the previous models done in Spatial Interaction Models. Will examining the variables further:

|                          |         |        |            |           |                                                                                  |
|---------------|---------|----------|------------|-------------|---------------|
|                          | **est** | **sd** | **t.stat** | **p.val** | **Remarks**                                                                      |
| rho_d                    | 0.677   | 0.001  | 565.502    | 0.000     |                                                                                  |
| rho_o                    | 0.663   | 0.001  | 543.029    | 0.000     |                                                                                  |
| rho_w                    | -0.449  | 0.001  | -407.064   | 0.000     |                                                                                  |
| (Intercept)              | 0.806   | 0.013  | 61.639     | 0.000     |                                                                                  |
| (Intra)                  | -0.723  | 0.026  | -27.305    | 0.000     |                                                                                  |
| D_SCHOOL_COUNT           | 0.001   | 0.001  | 1.074      | 0.283     | Surprisingly - this is not statistically significant.                            |
| D_SCHOOL_COUNT.lag1      | 0.007   | 0.003  | 2.602      | 0.009     |                                                                                  |
| D_BIZ_COUNT              | 0.000   | 0.000  | 0.923      | 0.356     | Surprisingly - this is not statistically significant.                            |
| D_BIZ_COUNT.lag1         | 0.000   | 0.000  | 2.869      | 0.004     |                                                                                  |
| D_FINSERV_COUNT          | 0.004   | 0.000  | 33.258     | 0.000     |                                                                                  |
| D_FINSERV_COUNT.lag1     | -0.001  | 0.000  | -8.567     | 0.000     |                                                                                  |
| D_n_busstops             | 0.010   | 0.000  | 34.871     | 0.000     |                                                                                  |
| D_n_busstops.lag1        | -0.009  | 0.001  | -17.052    | 0.000     |                                                                                  |
| O_n_busstops             | 0.009   | 0.000  | 31.563     | 0.000     |                                                                                  |
| O_n_busstops.lag1        | -0.008  | 0.001  | -14.450    | 0.000     |                                                                                  |
| O_TRAIN_EXITS_COUNT      | 0.014   | 0.001  | 22.403     | 0.000     |                                                                                  |
| O_TRAIN_EXITS_COUNT.lag1 | -0.017  | 0.001  | -14.768    | 0.000     |                                                                                  |
| O_FOOD_COUNT             | 0.000   | 0.000  | 2.437      | 0.015     | Surprisingly - this is statistically significant. But the t.stat is rather low.  |
| O_FOOD_COUNT.lag1        | 0.001   | 0.000  | 4.038      | 0.000     |                                                                                  |
| O_POP                    | 0.000   | 0.000  | 4.374      | 0.000     |                                                                                  |
| O_POP.lag1               | 0.000   | 0.000  | -0.637     | 0.524     |                                                                                  |
| O_CHILDCARE_COUNT        | 0.003   | 0.001  | 5.961      | 0.000     | Surprisingly - this is statistically significant. But the t.stat is rather low.  |
| O_CHILDCARE_COUNT.lag1   | 0.001   | 0.001  | 0.946      | 0.344     |                                                                                  |
| P_log(DISTANCE + 1)      | -0.086  | 0.001  | -65.968    | 0.000     |                                                                                  |

It sseems to suggest that travel is still very much explained by the number of bus stops within that area, the number of train station exits at that area and that there are certain type of industry, such as FINSERV that will see most of the people travelling to by bus etc.

Looking at the multi-collinearity:

```{r}
corr_residual <- pair_cor(mle_model8)
colnames(corr_residual) <- substr(colnames(corr_residual),1,3)
cor_image(corr_residual, cex = 0.5)
```

```{r}
old_par <- par(mfrow = c(1, 3), 
               mar = c(2,2,2,2))
spflow_moran_plots(mle_model8)
```

# Conclusion

It seems that the Spatial Econometric Interaction Modelling did a better job in fitting the data. However, i would still need to further examine it and look at the other models as well. Submitting this portion first due to time constraint.

Nonetheless - I think I have learnt a lot during this course, not just about geospatial analytics, but also about the various tools that we can used to get data. It is certainly useful for us in the future.

Would like to thank Prof Kam for putting in so much effort in this short five weeks.
